What is the nature of decision noise in random exploration?

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/draft/Older drafts/Bob#1/Internal-external noise Bob#1/Internal-External Noise.tex
add older version

	Imagine trying to decide where to go to dinner with a friend. You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing. Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.

@@

	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown. Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain. In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants. Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Previous work makes a strong case for both types of noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{A key additional manipulation here is repeated games. Each pair of repeated games with identical example trials will appear twice during the experiment. We setup the repeated games such that they are at least 5 games apart from each other. A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree}
		\end{center}
	\end{figure}

@@

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven. In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely external and purely internal noise cases.  For purely external noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case. For purely internal noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
	\begin{equation*}
	\begin{split}
	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean})(1-p(\mbox{low mean}))
	\end{split}
	\end{equation*}
	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(59) = 5.10, p $<$ 0.001 for horizon 1, t(59) = 5.65, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 3.49, p $<$ 0.001 for horizon 1, t(59) = 3.31, p $<$ 0.001 for horizon 6. Behavior is different from pure external noise prediction in both [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency. Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.

@@

	Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:mb1}. Consistent with our model-free results, we see that both internal and external noise variances are non-zero and that internal noise is about 2-3 times larger than the external noise. In addition, we find that internal noise increases dramatically with horizon (M = 4.55, p $<$ 0.001) whereas there is only a slight increase of external noise (M = 1.78, p = 0.009).  Taken together, these results suggest that random exploration is dominated by internal noise.

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb1}
		\end{center}
	\end{figure}

@@

%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
%			\caption[]{}			
%			\label{fig:mb2}
%		\end{center}
%	\end{figure}

@@

	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being internal and external. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose. Thus, our estimate of external noise is likely a lower bound. Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of external noise would be interpreted as internal noise in our model. Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb1}, where internal noise is 2-3 times the size of external noise.

@@

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise. Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

@@

	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment. 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 60 for the main analysis. Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree2}
		\end{center}
	\end{figure}

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf22}
		\end{center}
	\end{figure}
	\newpage
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb12}
		\end{center}
	\end{figure}
	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities. (\ref{appendix:participants})

@@

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?  I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME

@@

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game. For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome. From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 300ms. The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}. (See Figure \ref{fig:taskfigb})

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
			\caption{Time line of a horizon 6 game.  The game begins with four forced choice trials in which participants are instructed which option to play.   After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
			I WILL SEND YOU THE POWERPOINTS TOO}
			\label{fig:taskfigb}
		\end{center}
	\end{figure}

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions). (See Figure \ref{fig:taskfigb})

%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
%			\caption{Task conditions}
%			\label{fig:taskfig2}
%		\end{center}
%	\end{figure}

@@

	\begin{table}[h]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			\hline
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
			\hline
			spatial bias, $b_{is}$ 
			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
			& \specialcell{
				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
			\hline
			deviation of external noise, $\sigma^{ext}_{isg}$ 
			& $\sigma_{i} \sim $  Gamma($k_i^{ext}$, $\lambda_{i}^{ext}$) 
			& $\theta_{i}^{ext} = (k_i^{ext}, \lambda_i^{ext}) $
			& \specialcell{
				$k_i^{ext} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
			\hline
			deviation of internal noise, $\sigma^{int}_{isgr}$ 
			& $\sigma_{is} \sim $  Gamma($k_i^{int}$, $\lambda_{i}^{int}$) 
			& $\theta_{i}^{int} = (k_i^{int}, \lambda_i^{int}) $
			& \specialcell{
				$k_i^{int} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{int} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
		\label{tab:pars2}	
	\end{table}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
			\caption{Hierarchical Bayesian model}
			\label{fig:model}
		\end{center}
	\end{figure}

@@

	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}. As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C, D).  The recovery for the noise parameters, $\sigma_{ext}$ and $\sigma_{int}$, is slightly better for horizon 1 than horizon 6. This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6. In addition we see better recovery for internal noise than external noise because we effectively have half as many trials for external noise since we are only generating one sample of external noise for each repeated game pair. Overall, we are able to recover both external and internal noises using our model to a satisfactory extent.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/precover.png}
			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.  REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.  DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.  RED LINES CHANGE TO DOTTED BLACK LINES.  
			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
			%USE ADDABCS FUNCTION TO ADD ABCs!
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
	%\begin{figure}[H]-
	%	\begin{center}
	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
	%		\label{fig:pararecover2}
	%	\end{center}
	%\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.

>>>>

	Imagine trying to decide where to go to dinner with a friend. You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing. Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is incredibly hard \citep{eegittins74} (CITE BELMAN). In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eebridle90} CITE THOMPSON AND NEW PAPERS ON THOMPSON SAMPLING.

@@

	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown. Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain. In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants. Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Previous work makes a strong case for both types of noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures CITES.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfig.PNG}
			\caption{A key additional manipulation here is repeated games. Each pair of repeated games with identical example trials will appear twice during the experiment. We setup the repeated games such that they are at least 5 games apart from each other. A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions. TAKE OUT PANEL C }
			\label{fig:modelfree}
		\end{center}
	\end{figure}

@@

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven. In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (STATS ANOVA WITH HORIZON AND UNCERTAINTY CONDITIONS AS FACTORS), suggesting that at least some of the horizon dependent noise is internal.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/modelfree_intext.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely external and purely internal noise cases.  For purely external noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case. For purely internal noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
	\begin{equation*}
	\begin{split}
	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean}) CHECK????
	\end{split}
	\end{equation*}
	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (STATS T-TEST RELATIVE TO BOTH), suggesting that both external and internal noise are present in driving this choice inconsistency. Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.

@@

	Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:mb1}. Consistent with our model-free results, we see that both internal and external noise variances are non-zero and that internal noise is about 2-3 times larger than the external noise. In addition, we find that internal noise increases dramatically with horizon (STATS) whereas there is only a slight increase of external noise (STATS).  Taken together, these results suggest that random exploration is dominated by internal noise.

@@

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoiseh.png}
			\caption[Posterior distributions over the group-level means of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise variance. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'}
			\label{fig:mb1}
		\end{center}
	\end{figure}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
			\caption[Posterior distributions over the group-level means of the change of external and internal noise variance.]{Posterior distributions over the group-level means of the change of external and internal noise variance. Internal noise increases significantly with horizon, external noise increases with horizon as well.}			
			\label{fig:mb2}
		\end{center}
	\end{figure}

@@

	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being internal and external. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose. Thus, our estimate of external noise is likely a lower bound. Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of external noise would be interpreted as internal noise in our model. Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb2}, where internal noise is 2-3 times the size of external noise.

@@

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise. Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks (CITES KARPOVA, PAPER I JUST SENT YOU AND REFS WITHIN, ALSO JANE'S BIOARXIV PAPER).  In addition there is some evidence that NE plays a direct role in random exploration (CITE WARREN ET AL. 2017), although this finding is complicated by other work showing no effect of NE drugs on exploration CITE NIEWENHUIS AND JEPMA DRUG STUDY.

@@

	80 participants (ages 18-XXX, XXX male, XXX female, XXX no response) from the University of Arizona undergraduate subject pool participated in the experiment.  20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 60 for the main analysis. Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures XXX-YYY). XXX MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities. (\ref{appendix:participants})

@@

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 10.6 and 29.2 minutes (mean 19.7 minutes).  IS THIS NUMBER CORRECT?  I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})

@@

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game. For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome. From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 300ms. The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}. (See Figure \ref{fig:taskfig1})

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfig1.PNG}
			\caption{Time line of a horizon 6 game.  The game begins with four forced choice trials in which participants are instructed which option to play.   After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
			COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
			CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
			\label{fig:taskfig1}
		\end{center}
	\end{figure}

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions). (See Figure \ref{fig:taskfig2})

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
			\caption{Task conditions}
			\label{fig:taskfig2}
		\end{center}
	\end{figure}

@@

	\begin{table}[h]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			\hline
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.1)}		\\
			\hline
			spatial bias, $B_{is}$ 
			& $B_{is} \sim $  Gaussian($\mu_i^{B}$, $\sigma_i^{B}$) 
			& $\theta_{i}^{B} = (\mu_i^{B}, \sigma_i^{B}) $
			& \specialcell{
				$\mu_i^{B} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{B} \sim $ Exponential(0.1)}		\\
			\hline
			deviation of external noise, $\epsilon_{isg}$ 
			& $\sigma_{i} \sim $  Gamma($k_i^{ext}$, $\lambda_{i}^{ext}$) 
			& $\theta_{i}^{ext} = (k_i^{ext}, \lambda_i^{ext}) $
			& \specialcell{
				$k_i^{ext} \sim $ Gaussian(0.1) \\ 
				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
			\hline
			deviation of internal noise, $\sigma_{isgr}$ 
			& $\sigma_{is} \sim $  Gamma($k_i^{int}$, $\lambda_{i}^{int}$) 
			& $\theta_{i}^{int} = (k_i^{int}, \lambda_i^{int}) $
			& \specialcell{
				$k_i^{int} \sim $ Gaussian(0.1) \\ 
				$\lambda_i^{int} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors. CHECK UPDATE}
		\label{tab:pars2}	
	\end{table}

@@

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/hbm.pdf}
			\caption{Hierarchical Bayesian model}
			\label{fig:model}
		\end{center}
	\end{figure}

@@

	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}. As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure XXX PANEL NUMBER).  The recovery for the noise parameters, $\sigma_{ext}$ and $\sigma_{int}$, is slightly better for horizon 1 than horizon 6. This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6. In addition we see better recovery for internal noise than external noise because we effectively have half as many trials for external noise since we are only generating one sample of external noise for each repeated game pair. Overall, we are able to recover both external and internal noises using our model to a satisfactory extent.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/paramrecover_1.png}
			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 1 games
			COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.  REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.  DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.  RED LINES CHANGE TO DOTTED BLACK LINES.  
			XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
			USE ADDABCS FUNCTION TO ADD ABCs!
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
			\label{fig:pararecover2}
		\end{center}
	\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.
